A Study of Institutional Spending on Open Access Publication Fees in Germany

### 2016-06-05 article.Rmd
abstract + conclusion

---
title: 'Disclosing funding sources for open access publication fees: the Open APC
  initiative'
author:
- Najko Jahn, Bielefeld University Library, Bielefeld University (TIB) - German National
  Library of Science and Technology, Hannover, Germany
- <a href="http://orcid.org/0000-0002-5111-2788">Marco Tullney</a>, Technische Informationsbibliothek
date: "7 April 2016"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: peerj.sty
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: no
    theme: readable
bibliography: literatur.bib
csl: frontiers.csl
abstract: Publication fees in open access publishing hold a prominent place on the
  agenda of researchers, policy-makers, and academic publishers. This paper contributes
  to the evolving empirical basis on open access funding. It describes the Open APC
  initiative, in which German universities and research organizations share their
  expenditures for publication fees. As method, the initiative uses existing open
  data tools to aggregate and disseminate institutional spending on open access publication
  fees. In total, 29 German research organizations self-reported funding of 6,279
  open access journal articles, which amounted to 8,039,339 €. The average payment
  for each article was 1,280 €, and the median payment 1,209 €. Our data-set comprises
  only 53 articles in hybrid journals. With an indexing coverage of 99 %, the findings
  reveal that the DOI agency Crossref provides both comprehensive bibliographic coverage
  of the funded open access journal literature and disambiguated names of journal
  titles and publishing houses. We show that authority control of these bibliographic
  information is particularly relevant for the comparative study of the economical
  effects of open access publishing.
---

@@

<!-- Publication fees, often paid by funders or universities, are a widely discussed business model for open access journals. Yet, how and to what extent these activities are effective in terms of the number of supported research articles and associated costs remains under debate. This study contributes to a growing empirical base about institutional spending on open access publication fees. More specifically, it uses self-reported cost data from German universities and research organisations to examine how much they spent, and how the so funded open access journal articles were distributed over publishers and journal titles using bibliographic information from Crossref. Such an approach extends methods and improves data collection activities for researchers and practitioners alike, as well as contribute to a better understanding of factors affecting the analysis of publication fees in open access publishing. -->

>>>>

---
title: 'A Study of Institutional Spending on Open Access Publication Fees in Germany'
author:
- Najko Jahn, Bielefeld University Library, Bielefeld University (TIB) - German National
  Library of Science and Technology, Hannover, Germany
- <a href="http://orcid.org/0000-0002-5111-2788">Marco Tullney</a>, Technische Informationsbibliothek
date: "7 April 2016"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: peerj.sty
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: no
    theme: readable
bibliography: literatur.bib
csl: frontiers.csl

---

@@
